---
title: "Lab 1"
author: "Nico Wisebaker"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

Lab 1 Tasks

## Task 1

```{r}
getwd()
```


## Task 2
```{r}
ddt <- read.csv("DDT.csv")
head(ddt)
```


## Task 3
```{r}
table(ddt$SPECIES)

ddt[ddt$SPECIES == "LMBASS" & ddt$WEIGHT > 800,]

ddt[ddt$RIVER == "SCM" & ddt$DDT > 4.0,]
```
There are 3 species.
Two LMBASS have a weight over 800.


## Task 4
```{r}
length <- ddt$LENGTH
mean(length)
```

Q1: 42.8125
```{r}
weight <- ddt$WEIGHT
sd(weight)
```

Q2: 376.5461
```{r}
with(ddt,plot(WEIGHT,LENGTH))
```

Q3: 0
```{r}
v = 1:20
v/20
```

Q4: 1.00


## Task 5
```{r}
tbl1 <- with(ddt,table(RIVER))
tbl1
```

Table of RIVER
```{r}
colorsYay <- c("#faad11", "#fe679a", "#a811fc", "#6c2afb")
barplot(tbl1,col = colorsYay)
```

Bar plot of RIVER (color-coded)
```{r}
tbl2 <- with(ddt,table(RIVER,SPECIES))
tbl2
```

Table of RIVERS crossed with SPECIES
```{r}
barplot(tbl2, col = colorsYay)
```

Bar plot of RIVERS crossed with SPECIES


## Task 6
```{r}
pie1 <- with(ddt, table(SPECIES))
pie(pie1, col = colorsYay)
```

Pie chart of SPECIES
```{r}
pie2 <- with(ddt, table(RIVER))
pie(pie2, col = colorsYay)
```

Pie chart of RIVER


## Task 7
```{r}
ddtBox <- with(ddt, boxplot(DDT, ylab="DDT", col = colorsYay[2], notch=TRUE))
```

Box plot of the variable DDT
```{r}
weightBox <- with(ddt,boxplot(WEIGHT, ylab="WEIGHT", col = colorsYay[3], notch=TRUE))
```

Box plot of the variable WEIGHT
```{r}
lengthBox <- with(ddt,boxplot(LENGTH, ylab="LENGTH", col = colorsYay[4], notch=TRUE))
```

Box plot of the variable LENGTH


## Task 8
```{r}
lengthWeight <- with(ddt, coplot(LENGTH~WEIGHT|RIVER, col = colorsYay))
```

Coplot of LENGTH vs WEIGHT given RIVER
```{r}
m=with(ddt, as.numeric(factor(MILE))) # A
m
length(unique(m)) #B
ddtWeight <- with(ddt, coplot(LENGTH~WEIGHT|RIVER*SPECIES, col = MILE))
```

Coplot of LENGTH Vs WEIGHT given RIVER*SPECIES


## Task 9
```{r}
library(ggplot2)
g1 <- ggplot(data = ddt, aes(x = SPECIES, y = WEIGHT, fill  = RIVER))
g1 <- g1 + geom_boxplot() + ggtitle("Nico Wisebaker's Box Plot") +scale_fill_manual(values = colorsYay)
g1
g2 <- ggplot(data = ddt, aes(x = RIVER, y = LENGTH, fill  = SPECIES))
g2 <- g2 + geom_violin() + ggtitle("Nico Wisebaker's Violin Plot") + scale_fill_manual(values = colorsYay)
g2
g3 <- ggplot(data = ddt, aes(x = WEIGHT, y = LENGTH, col = SPECIES))
g3 <- g3 + geom_point() + ggtitle("Nico Wisebaker's Scatter Plot") + scale_color_manual(values = colorsYay)
g3
```

