---
title: "Lab 6 Markdown"
author: "Nico Wisebaker"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1
```{r}
getwd()
```

## Task 2

```{r}
layout(matrix(1:4, nr=2,nc=2))
layout.show(4)

curve(dnorm(x, mean=10,sd=4),xlim=c(-10,30))

curve(dnorm(x, mean=10,sd=2),xlim=c(-10,30))

curve(dnorm(x, mean=5,sd=10),xlim=c(-40,50))

curve(dnorm(x, mean=5,sd=0.5),xlim=c(0,10))
```

Four plots
```{r}
curve(dnorm(x, mean=0,sd=1),xlim=c(-5,5)) # P(Y >=2)
xcurve=seq(2,5,length=1000)
ycurve=dnorm(xcurve,mean=0,sd=1)
polygon(c(2,xcurve,5),c(0,ycurve,0),col="Red")
prob=pnorm(5,mean=0,sd=1)-pnorm(2,mean=0,sd=1)
prob=round(prob,4)
text(x=2.2, y=0.2, paste("Area = ", prob, sep=""))

curve(dnorm(x, mean=4,sd=2),xlim=c(-3,11)) # P(1 <= Y < 5)
xcurve=seq(1,5,length=1000)
ycurve=dnorm(xcurve,mean=4,sd=2)
polygon(c(1,xcurve,5),c(0,ycurve,0),col="Red")
prob=pnorm(5,mean=4,sd=2)-pnorm(1,mean=4,sd=2)
prob=round(prob,4)
text(x=8, y=0.1, paste("Area = ", prob, sep=""))

curve(dnorm(x, mean=10,sd=4),xlim=c(-10,30)) # P(Y < 10)
xcurve=seq(-10,10,length=1000)
ycurve=dnorm(xcurve,mean=10,sd=4)
polygon(c(-10,xcurve,10),c(0,ycurve,0),col="Red")
prob=pnorm(10,mean=10,sd=4)-pnorm(-10,mean=10,sd=4)
prob=round(prob,4)
text(x=20, y=0.04, paste("Area = ", prob, sep=""))


curve(dnorm(x, mean=-2,sd=0.5),xlim=c(-5,1)) # P(-3 < Y <= -2)
xcurve=seq(-3,-2,length=1000)
ycurve=dnorm(xcurve,mean=-2,sd=0.5)
polygon(c(-3,xcurve,-2),c(0,ycurve,0),col="Red")
prob=pnorm(-2,mean=-2,sd=0.5)-pnorm(-3,mean=-2,sd=0.5)
prob=round(prob,4)
text(x=-0.5, y=0.4, paste("Area = ", prob, sep=""))

```

Four separate plots

## Task 3
```{r}
curve(dgamma(x,shape=1,scale=1),xlim=c(0,10),ylim=c(0,1),col="Red",lwd=2,
ylab="Gamma density", main="Beta=1")
text(x=1.2,y=0.7, paste("alpha=",1), col = "red")

curve(dgamma(x,shape=3,scale=1),xlim=c(0,10),ylim=c(0,1),add=TRUE,lwd=2)
text(x=1.8,y=0.4, paste("alpha=",3))

curve(dgamma(x,shape=5,scale=1),xlim=c(0,10),ylim=c(0,1),add=TRUE, col="Blue",lwd=2)
text(x=3.8,y=0.3, paste("alpha=",5), col = "blue")
```

3 plots together

```{r}
curve(dgamma(x,shape=3,scale=2),xlim=c(0,20)) # P(2<Y<5)
xcurve=seq(2,5,length=1000)
ycurve=dgamma(xcurve,shape=3,scale=2)
polygon(c(2,xcurve,5),c(0,ycurve,0),col="Red")
prob=pgamma(5,shape=3,scale=2)-pgamma(2,shape=3,scale=2)
prob=round(prob,4)
text(x=10, y=0.1, paste("Area = ", prob, sep=""))

curve(dgamma(x,shape=6,scale=3),xlim=c(0,50)) # P(1≤Y≤4)
xcurve=seq(1,4,length=1000)
ycurve=dgamma(xcurve,shape=6,scale=3)
polygon(c(1,xcurve,4),c(0,ycurve,0),col="Red")
prob=pgamma(4,shape=6,scale=3)-pgamma(1,shape=6,scale=3)
prob=round(prob,4)
text(x=3, y=0.01, paste("Area = ", prob, sep=""))

curve(dgamma(x,shape=2,scale=4),xlim=c(0,30)) # P(3≤Y<6)
xcurve=seq(3,6,length=1000)
ycurve=dgamma(xcurve,shape=2,scale=4)
polygon(c(3,xcurve,6),c(0,ycurve,0),col="Red")
prob=pgamma(6,shape=2,scale=4)-pgamma(3,shape=2,scale=4)
prob=round(prob,4)
text(x=11, y=0.08, paste("Area = ", prob, sep=""))

```
The 3 separate area graphs

## Task 4
```{r}
layout(matrix(1:4, nr=2,nc=2))
layout.show(4)

curve(dchisq(x,df=1),xlim=c(0,5))

curve(dchisq(x,df=2),xlim=c(0,8))

curve(dchisq(x,df=4),xlim=c(0,15))

curve(dchisq(x,df=20),xlim=c(0,50))
```
Four chsq graphs

```{r}
curve(dchisq(x,df=2),xlim=c(0,8)) # P(2≤Y≤4)
xcurve=seq(2,4,length=1000)
ycurve=dchisq(xcurve,df=2)
polygon(c(2,xcurve,4),c(0,ycurve,0),col="Red")
prob=pchisq(4,df=2)-pchisq(2,df=2)
prob=round(prob,4)
text(x=3, y=0.2, paste("Area = ", prob, sep=""))

curve(dchisq(x,df=3),xlim=c(0,10)) # P(3≤Y≤5)
xcurve=seq(3,5,length=1000)
ycurve=dchisq(xcurve,df=3)
polygon(c(3,xcurve,5),c(0,ycurve,0),col="Red")
prob=pchisq(5,df=3)-pchisq(3,df=3)
prob=round(prob,4)
text(x=4.5, y=0.15, paste("Area = ", prob, sep=""))

curve(dchisq(x,df=20),xlim=c(0,50)) # P(10<Y≤21)
xcurve=seq(10,21,length=1000)
ycurve=dchisq(xcurve,df=20)
polygon(c(10,xcurve,21),c(0,ycurve,0),col="Red")
prob=pchisq(21,df=20)-pchisq(10,df=20)
prob=round(prob,4)
text(x=5, y=0.03, paste("Area = ", prob, sep=""))
```
3 chisq area graphs graphed separately

## Task 5
```{r}
layout(matrix(1:4, nr=2,nc=2))
layout.show(4)

curve(pweibull(x, shape=1, scale=3), xlim=c(0,10))

curve(dweibull(x, shape=8, scale=4), xlim=c(0,8))

curve(pweibull(x, shape=9999, scale=2), xlim=c(0,4))

curve(dweibull(x, shape=5, scale=7), xlim=c(0,12))
```

Four graphs of choice

```{r}
aaaa = curve(pweibull(x, shape=1, scale=3), xlim=c(0,10)) # P(Y > 0) -- sorry
xcurve=seq(0,length=1000)
ycurve=pweibull(xcurve,shape=1,scale=3)
polygon(c(0,xcurve,10),c(0,ycurve,0),col="Red")
prob=pweibull(xcurve, shape=1, scale=3)
prob=round(prob,4)
text(x=1, y=0.6, paste("Area = ", prob, sep=""))

curve(dweibull(x, shape=8, scale=4), xlim=c(0,8)) # P( Y < 4)
xcurve=seq(0,4,length=1000)
ycurve=dweibull(xcurve,shape=8,scale=4)
polygon(c(0,xcurve,4),c(0,ycurve,0),col="Red")
prob=dweibull(4, shape=8, scale=4)-dweibull(0, shape=8, scale=4)
prob=round(prob,4)
text(x=2.5, y=0.5, paste("Area = ", prob, sep=""))

curve(pweibull(x, shape=9999, scale=2), xlim=c(0,4)) # P( 1 < Y < 3)
xcurve=seq(1,3,length=1000)
ycurve=pweibull(xcurve,shape=9999,scale=2)
polygon(c(1,xcurve,3),c(0,ycurve,0),col="Red")
prob=pweibull(3, shape=9999, scale=2)-pweibull(1, shape=9999, scale=2)
prob=round(prob,4)
text(x=2.5, y=0.5, paste("Area = ", prob, sep=""))
```

Three areas of probabilities of my choosing, graphed separately, I couldn't quite get P(Y > 0) to be 1 though :(... I'm also not sure about the last one.

## Task 6
```{r}
layout(matrix(1:4, nr=2,nc=2))
layout.show(4)

curve(dbeta(x, shape1=1, shape2=3), xlim=c(0,3))

curve(dbeta(x, shape1=8, shape2=4), xlim=c(0,1.5))

curve(dbeta(x, shape1=9999, shape2=2), xlim=c(0,2))

curve(dbeta(x, shape1=5, shape2=7), xlim=c(0,1))
```
Four dbeta curves

```{r}
curve(dbeta(x, shape1=1, shape2=3), xlim=c(0,3)) # P(Y>0.5)
xcurve=seq(0,0.5,length=1000)
ycurve=dbeta(xcurve,shape1=1,shape2=3)
polygon(c(0,xcurve,0.5),c(0,ycurve,0),col="Red")
prob=pbeta(0.5, shape1=1, shape2=3)
prob=round(prob,4)
text(x=0.5, y=1.5, paste("Area = ", prob, sep=""))

curve(dbeta(x, shape1=8, shape2=4), xlim=c(0,1.5)) # P( 0.7 < Y < 0.9)
xcurve=seq(0.7,0.9,length=1000)
ycurve=dbeta(xcurve,shape1=8,shape2=4)
polygon(c(0.7,xcurve,0.9),c(0,ycurve,0),col="Red")
prob=pbeta(0.9, shape1=8, shape2=4)
prob=round(prob,4)
text(x=1, y=1.5, paste("Area = ", prob, sep=""))

curve(dbeta(x, shape1=9999, shape2=2), xlim=c(0,2)) # P( Y < 1)
xcurve=seq(0,0.9999,length=1000)
ycurve=dbeta(xcurve,shape1=9999,shape2=2)
polygon(c(0,xcurve,0.9999),c(0,ycurve,0),col="Red")
prob=pbeta(0.9999, shape1=9999, shape2=2)
prob=round(prob,4)
text(x=1, y=0, paste("Area = ", prob, sep=""))
```
Three separate dbeta areas of my choosing

## Task 7

Blank :(

## Task 8

Blank :(