---
title: "Lab 2"
author: "Nico Wisebaker"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Tasks

Lab 2 Tasks

## Task 1

```{r}
getwd()
```

## Task 2
```{r}
EPAGAS <- read.csv("EPAGAS.csv")
head(EPAGAS)
```

Read EPAGAS.csv and obtained the first six lines of data

## Task 3

```{r}
mpg <- EPAGAS$MPG
```

Created the mpg vector

```{r}
xBar <- mean(mpg)
sdX <- sd(mpg)

z <- (mpg - xBar) / sdX
```

Transformed mpg to z-scores

```{r}
sdZ <- sd(z)
possibleOutliers <- mpg[abs(z) > 2 * sdZ]
possibleOutliers
```

Values of mpg that are possible outliers (i.e., z-scores that are beyond 2 standard deviations from the mean)

```{r}
outliers <- mpg[abs(z) > 3 * sdZ]
outliers
```

Values of mpg that are defined as outliers (3 sd away).

```{r}
library(lattice)
outlierLabels <- rep("Normal", length(mpg))
outlierLabels[mpg %in% outliers] <- "Outlier"
outlierLabels[mpg %in% possibleOutliers] <- "Possible Outlier"
mycol = ifelse(abs(z)>3, "Red",
        ifelse(abs(z)>=2 &abs(z)<=3,"Blue", "Black"))
dotplot(mpg,col=mycol,)
```

Constructed a dotplot with colors Red=outliers & Blue=possible outliers

## Task 4

```{r}
boxplot(mpg, horizontal = TRUE, col = "black", notch = TRUE, border = "#343434", main = "Nico Wisebaker's Boxplot of MPG")
```
Made a boxplot of the mpg variable

```{r}
chebyshevProp <- 1 - 1/2^2
library(scales)
label_percent()(chebyshevProp)
```
Predicted the proportion of data within 2 standard deviations of the mean of the mpg data using Chebyshev's thm.

```{r}
within2sd <- mean(abs(mpg - xBar) <= 2 * sdX)
label_percent()(within2sd)
```
Calculated the proportion of data within 2 standard deviations of the mean of the mpg data.

According to Chebyshev's theorem, at least 75% of the data should fall within 2 standard deviations of the mean, and it holds up here. Although this lower bound is not very close to the actual result.

According to the emperical rule, this proportion should be 95%. This dataset corresponds fairly well with the emperical rule, as the prediction for the proportion of data within 2 standard deviations is only a 1% difference. So, the data is approximately normal.

Checking validity:

```{r}
hist(mpg)
```

The emperical rule is valid since the data is unimodal and symmetric about the mean.