---
title: "Lab 8 Markdown"
author: "Nico Wisebaker"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1

```{r}
getwd()
```

# Task 2

```{r}
a = 0
b = 5
runif(10,a,b)
```

```{r}
sample = c(1.3203239, 4.8155007, 0.6193721, 0.5776914, 4.7227894, 4.1488387, 3.9591730,
           1.7561037, 3.436954, 3.6542493)
cat("sample:", sample, "\n")
mu = (a+b)/2
sigmasq = ((b-a)^2)/12
sigma = sqrt(sigmasq)
cat("population mean:", mu, "\n")
cat("population variance:", sigmasq, "\n")
xbar = mean(sample)
ssq = var(sample)
cat("sample mean:", xbar, "\n")
cat("sample variance:", ssq, "\n")
```
With this particular random sample, my sample mean was higher than the population mean by about 0.4 and my sample variance was higher than the population variance by about 0.7

```{r}
E_T = 10*mu
V_T = 10*sigmasq
E_x = mu
V_x = sigmasq/10
cat("sum mean:", E_T, "\n")
cat("sum variance:", V_T, "\n")
cat("mean mean:", E_x, "\n")
cat("mean variance:",V_x, "\n")
```
```{r}
myclt=function(n,iter){
y=runif(n*iter,0,5) # A
data=matrix(y,nr=n,nc=iter,byrow=TRUE) #B
sm=apply(data,2,sum) #C
hist(sm)
sm
}
w=myclt(n=10,iter=10000) #D

mean(w)
var(w)
```
Line A creates y, a random uniform distribution of n*iter variables, with a minimum of 0 and a maximum of 5.
Line B creates data, a matrix from the data in y with n rows and iter columns. 
Line C creates sm, the sum of the numbers in the columns of the data matrix.
Line D uses the function myclt with n=10 and iter=10000

```{r}
myclt=function(n,iter){
y=runif(n*iter,0,5) # A
data=matrix(y,nr=n,nc=iter,byrow=TRUE) #B
sm=apply(data,2,sum) #C
hist(sm/n)
sm/n
}
w=myclt(n=10,iter=10000) #D
mean(w)
var(w)
```

# Task 3

```{r}
mycltu=function(n,iter,a=0,b=10){
## r-random sample from the uniform
y=runif(n*iter,a,b)
## Place these numbers into a matrix
## The columns will correspond to the iteration and the rows will equal the sample size n
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
## apply the function mean to the columns (2) of the matrix
## these are placed in a vector w
w=apply(data,2,mean)
## We will make a histogram of the values in w
## How high should we make y axis?
## All the values used to make a histogram are placed in param (nothing is plotted yet)
param=hist(w,plot=FALSE)
## Since the histogram will be a density plot we will find the max density

ymax=max(param$density)
## To be on the safe side we will add 10% more to this
ymax=1.1*ymax
## Now we can make the histogram
hist(w,freq=FALSE,  ylim=c(0,ymax), main=paste("Histogram of sample mean",
"\n", "sample size= ",n,sep=""),xlab="Sample mean")
## add a density curve made from the sample distribution
lines(density(w),col="Blue",lwd=3) # add a density plot
## Add a theoretical normal curve 
curve(dnorm(x,mean=(a+b)/2,sd=(b-a)/(sqrt(12*n))),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
## Add the density from which the samples were taken
curve(dunif(x,a,b),add=TRUE,lwd=4)
}
mycltu(n=20,iter=100000)
```

-	`w=apply(data,2,mean)`, how does the apply function use the 2?
THe apply function uses the 2 to to find the mean of the columns (since the order is rows, columns) of the matrix

-	How many terms are in w, when `mycltu(n=20,iter=100000)` is called?
w should have 100000 terms since the matrix has 100000 columns

-	`curve(dnorm(x,mean=(a+b)/2`
creates a normal curve for the data

-	`sd=(b-a)/(sqrt(12*n))),add=TRUE,col="Red",lty=2,lwd=3)`: Explain why sd takes the formula as shown in the function.
Because of the formula for variance of means for a uniform distribution.

```{r}
mycltu(n=1, iter=10000, a=0,b=10)
mycltu(n=2, iter=10000, a=0,b=10)
mycltu(n=3,iter=10000,a=0,b=10)
mycltu(n=5,iter=10000,a=0,b=10)
mycltu(n=10, iter=10000,a=0,b=10)
mycltu(n=30,iter=10000,a=0,b=10)
```
As n increases, the distribution begins to look more like a normal distribution.

# Task 4

```{r}
mycltb=function(n,iter,p=0.5,...){

## r-random sample from the Binomial
y=rbinom(n*iter,size=n,prob=p)
## Place these numbers into a matrix
## The columns will correspond to the iteration and the rows will equal the sample size n
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
## apply the function mean to the columns (2) of the matrix
## these are placed in a vector w
w=apply(data,2,mean)
## We will make a histogram of the values in w
## How high should we make y axis?
## All the values used to make a histogram are placed in param (nothing is plotted yet)
param=hist(w,plot=FALSE)
## Since the histogram will be a density plot we will find the max density

ymax=max(param$density)
## To be on the safe side we will add 10% more to this
ymax=1.1*ymax

## Now we can make the histogram
## freq=FALSE means take a density
hist(w,freq=FALSE,  ylim=c(0,ymax),
main=paste("Histogram of sample mean","\n", "sample size= ",n,sep=""),
xlab="Sample mean",...)
## add a density curve made from the sample distribution
#lines(density(w),col="Blue",lwd=3) # add a density plot
## Add a theoretical normal curve 
curve(dnorm(x,mean=n*p,sd=sqrt(p*(1-p))),add=TRUE,col="Red",lty=2,lwd=3) 

}

mycltb(n=4,iter=10000,p=0.3)
mycltb(n=5,iter=10000,p=0.3)
mycltb(n=10,iter=10000,p=0.3)
mycltb(n=20,iter=10000,p=0.3)

mycltb(n=4,iter=10000,p=0.7)
mycltb(n=5,iter=10000,p=0.7)
mycltb(n=10,iter=10000,p=0.7)
mycltb(n=20,iter=10000,p=0.7)

mycltb(n=4,iter=10000,p=0.5)
mycltb(n=5,iter=10000,p=0.5)
mycltb(n=10,iter=10000,p=0.5)
mycltb(n=20,iter=10000,p=0.5)

```
I am not sure what to conclude, other than than the distributions seem consistent even if p changes. 


# Task 5

```{r}
mycltp=function(n,iter,lambda=10,...){

## r-random sample from the Poisson
y=rpois(n*iter,lambda=lambda)
## Place these numbers into a matrix
## The columns will correspond to the iteration and the rows will equal the sample size n
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
## apply the function mean to the columns (2) of the matrix
## these are placed in a vector w
w=apply(data,2,mean)
## We will make a histogram of the values in w
## How high should we make y axis?
## All the values used to make a histogram are placed in param (nothing is plotted yet)
param=hist(w,plot=FALSE)
## Since the histogram will be a density plot we will find the max density

ymax=max(param$density)
## To be on the safe side we will add 10% more to this
ymax=1.1*ymax

## Make a suitable layout for graphing
layout(matrix(c(1,1,2,3),nr=2,nc=2, byrow=TRUE))

## Now we can make the histogram
hist(w,freq=FALSE,  ylim=c(0,ymax), col=rainbow(max(w)),
main=paste("Histogram of sample mean","\n", "sample size= ",n," iter=",iter," lambda=",lambda,sep=""),
xlab="Sample mean",...)
## add a density curve made from the sample distribution
#lines(density(w),col="Blue",lwd=3) # add a density plot
## Add a theoretical normal curve 
curve(dnorm(x,mean=lambda,sd=sqrt(lambda/n)),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve

# Now make a new plot
# Since y is discrete we should use a barplot
barplot(table(y)/(n*iter),col=rainbow(max(y)), main="Barplot of sampled y", ylab ="Rel. Freq",xlab="y" )
x=0:max(y)
plot(x,dpois(x,lambda=lambda),type="h",lwd=5,col=rainbow(max(y)),
main="Probability function for Poisson", ylab="Probability",xlab="y")
}

mycltp(n=2, iter=10000,lambda=4)
mycltp(n=3,iter=10000,lambda=4)
mycltp(n=5,iter=10000,lambda=4)
mycltp(n=10,iter=10000,lambda=4)
mycltp(n=20,iter=10000,lambda=4)

mycltp(n=2, iter=10000,lambda=10)
mycltp(n=3,iter=10000,lambda=10)
mycltp(n=5,iter=10000,lambda=10)
mycltp(n=10,iter=10000,lambda=10)
mycltp(n=20,iter=10000,lambda=10)
```
Task video
<video width="320" height="240" controls>
  <source src="vid.mp4" type="video/mp4">
Your browser does not support the video tag.
</video>

# Task 6

```{r}
SPRING244753wise0073::myclt(1,2)
```

# Task 7 (Optional)

```{r}
```

